\documentclass[aspectratio=169,10pt]{beamer}

\usetheme{Madrid}
\usecolortheme{seahorse}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\usepackage{amsmath,amssymb,booktabs,array}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}
\usepackage{xcolor}
\usepackage{hyperref}

\definecolor{darkblue}{RGB}{0,65,120}
\hypersetup{colorlinks=true,linkcolor=darkblue,urlcolor=darkblue}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{TODO:}} \emph{#1}}

\title[L02: Distribution PF]{L02 -- Distribution Network Power Flow}
\subtitle{Foundations Lecture Series}
\author{Course Instructor}
\institute{Microgrid 3-phase Security AI Project}
\date{Foundations Lecture 2}

\begin{document}

\begin{frame}
  \titlepage
  \begin{center}
    \small Keywords: radial topology, R/X ratio, backward/forward sweep, DER, voltage rise
  \end{center}
\end{frame}

% ============================================================
\begin{frame}{Lecture roadmap}
\begin{enumerate}
  \item Why distribution differs from transmission
  \item Radial topology and the R/X problem
  \item Backward/Forward Sweep (BFS) algorithm
  \item DistFlow and linear voltage-drop approximations
  \item DER impact: reverse flow, voltage rise, hosting capacity
  \item Toward unbalanced analysis (preview of L04)
\end{enumerate}
\vspace{0.6em}
\begin{block}{Prerequisite}
L01 -- you should be comfortable with $Y_{\text{bus}}$, NR, and \code{pp.runpp}.
\end{block}
\end{frame}

% ============================================================
\section*{1. Why distribution differs}

\begin{frame}{Transmission vs distribution: a side-by-side}
\TODO{Build a comparison table covering: voltage level, topology (meshed vs radial), R/X ratio, load composition, three-phase symmetry, DER penetration. Highlight that distribution breaks several assumptions L01 relied on.}
\end{frame}

\begin{frame}{The R/X problem in one slide}
\TODO{Explain that high R/X ratio (typical for LV/MV feeders) makes the standard decoupled approximation invalid. Show the simplified voltage-drop formula and note that the P-term cannot be ignored. Set up motivation for BFS.}
\end{frame}

% ============================================================
\section*{2. Backward/Forward Sweep}

\begin{frame}{Backward/Forward Sweep at a glance}
\TODO{Algorithm sketch: traverse radial tree from leaves to root summing currents, then from root back to leaves updating voltages. Emphasize that it exploits the radial structure and avoids the Jacobian altogether.}
\end{frame}

\begin{frame}{When BFS beats Newton-Raphson}
\TODO{Convergence comparison on radial vs lightly-meshed networks. Note that BFS is sometimes the only solver that converges on very long LV feeders. pandapower exposes BFSW via \code{algorithm="bfsw"}.}
\end{frame}

% ============================================================
\section*{3. Linear approximations}

\begin{frame}{LinDistFlow and friends}
\TODO{Introduce the linearized voltage-drop formula and the DistFlow / LinDistFlow framework. Explain when the linearization is accurate and when it lies (high P injection, high R/X).}
\end{frame}

% ============================================================
\section*{4. DER on distribution feeders}

\begin{frame}{Reverse power flow and voltage rise}
\TODO{Show how a high-PV bus can have $V > 1.05$ pu (voltage rise) and how PV at the end of a feeder can cause reverse flow into the substation. Explain hosting capacity in one sentence.}
\end{frame}

\begin{frame}{The microgrid extension}
\TODO{Highlight grid-connected vs islanded modes, PCC role, DER + BESS as controllable resources. Forward-reference Week 1 kickoff material.}
\end{frame}

% ============================================================
\section*{5. Toward unbalanced analysis}

\begin{frame}{Why distribution is rarely balanced in practice}
\TODO{List sources of asymmetry: single-phase loads, single-phase rooftop PV, EV chargers, asymmetric line geometry. Conclude that L03's balanced assumption is convenient but not realistic for LV microgrids. Bridge to L04.}
\end{frame}

% ============================================================
\begin{frame}{Homework}
\TODO{Suggest: implement BFSW by hand on a 5-bus radial feeder; vary R/X ratio and compare to NR; analyze a feeder with one high-PV bus and observe voltage rise.}
\end{frame}

\begin{frame}{Exit checklist}
\TODO{Fill once content frames are finalized.}
\end{frame}

\begin{frame}{References}
\TODO{Baran \& Wu (DistFlow), Cespedes (1990 BFS paper), Kersting textbook for distribution analysis, pandapower BFSW documentation.}
\end{frame}

\end{document}
